Global spirals in self-gravitating dead zones in protoplanetary discs

### 2014-10-30 paper/results.tex
updated with fargo exchange simulations

\begin{figure}
  % scale=0.41
  \includegraphics[scale=.41]{figures/nonaxi_evol_ang_fargo}
  \caption{Evolution of angular momentum components in the fiducial
    FARGO simulation. The perturbation
    relative to $t=0$ in 2D is shown in units of the
    initial total angular momentum $J_\mathrm{ref}$.\label{2d_angmom}} 
\end{figure}

\subsection{Non-linear initial perturbation}
In order to focus on the one-arm spiral that emerges in the above
simulation, we here analyse a run with only an initial $m=1$
perturbation in the dead zone. Specifically, we set
\begin{align}
  &\Sigma \to \Sigma_0(R) + \delta\Sigma_0(R) + 2\Sigma_1(R)\cos{\phi},\\
  &v_\phi  \to v_{\phi 0}(R) + 2v_{\phi 1}(R)\cos{\phi},
\end{align}
for $R\in[R_{d1},R_{d2}]$, where
\begin{align}
  &\delta\Sigma_0 =
  -\delta^2\times\Sigma_0(R_{d1})\frac{R_{d1}}{R}\sin^3{\left[\frac{2\pi\left(R-R_{d1}\right)}{R_{d2}-R_{d1}}\right]},\\ 
  &\Sigma_1 =
  \delta\times\Sigma_0\sin{\left[\frac{\pi\left(R-R_{d1}\right)}{R_{d2}-R_{d1}}\right]},  
\end{align}
with $\delta = 10^{-3}$, and
\begin{align}
  v_{\phi 1} = -\frac{v_{\phi 0}\delta\Sigma_0}{2\Sigma_1}. 
\end{align}
This perturbs the dead zone structure while conserving total mass and
angular momentum. The $m=1$ perturbation has an associated negative
angular momentum (as observed in the previous simulation) at the
expense of increasing the angular momentum associated with the
background disc. %local exchange for simplicity 
%actually doesn't matter if we set plus or minus
Note that the perturbation to the axisymmetric surface density is
$O(\delta^2)$, whereas that in $m=1$ is $O(\delta)$. Since $Q>1$
everywhere in the disc, there is no risk of axisymmetric gravitational
instability.

@@

>>>>

\begin{figure}
  % scale=0.41
  \includegraphics[width=\linewidth]{figures/nonaxi_evol_ang_fargo}
  \caption{Evolution of angular momentum components in the fiducial
    FARGO simulation. The perturbation
    relative to $t=0$ in 2D is shown in units of the
    initial total angular momentum $J_\mathrm{ref}$.\label{2d_angmom}} 
\end{figure}

\subsection{Non-linear initial perturbation}\label{fargo_nonlin}
In order to focus on the one-arm spiral that emerges in the above
simulation, we here analyse a run with only an initial $m=1$
perturbation in the dead zone. Specifically, we set
\begin{align}
  &\Sigma \to \Sigma_0(R) + \delta\Sigma_0(R) + 2\Sigma_1(R)\cos{\phi},\\
  &v_\phi  \to v_{\phi 0}(R) + 2v_{\phi 1}(R)\cos{\phi},
\end{align}
for $R\in[R_{d1},R_{d2}]$, where
\begin{align}
  &\delta\Sigma_0 =
  -\delta^2\times\Sigma_0(R_{d1})\frac{R_{d1}}{R}\sin^3{\left[\frac{2\pi\left(R-R_{d1}\right)}{R_{d2}-R_{d1}}\right]},\\ 
  &\Sigma_1 =
  \delta\times\Sigma_0\sin{\left[\frac{\pi\left(R-R_{d1}\right)}{R_{d2}-R_{d1}}\right]},  
\end{align}
with $\delta = 10^{-3}$, and
\begin{align}
  v_{\phi 1} = -\frac{v_{\phi 0}\delta\Sigma_0}{2\Sigma_1}. 
\end{align}
This perturbs the dead zone structure while conserving total mass and
angular momentum. The $m=1$ perturbation has an associated negative
angular momentum (as observed in the previous simulation) at the
expense of increasing the angular momentum associated with the
background disc. %local exchange for simplicity 
%actually doesn't matter if we set plus or minus
Note that the perturbation to the axisymmetric surface density is
$O(\delta^2)$, whereas that in $m=1$ is $O(\delta)$. Since $Q>1$
everywhere in the disc, there is no risk of axisymmetric gravitational
instability.

@@

\begin{figure*}
  % scale=0.41
  \includegraphics[scale=0.4,clip=true,trim=0cm 0cm 1.73cm
  0cm]{figures/polarxy_fargo_ex_100}\includegraphics[scale=0.4,clip=true,trim=2.24cm
  0cm 1.73cm 0cm]{figures/polarxy_fargo_ex_150}\includegraphics[scale=0.4,clip=true,trim=2.24cm
  0cm 1.73cm 0cm]{figures/polarxy_fargo_ex_200}\includegraphics[scale=0.4,clip=true,trim=2.24cm
  0cm 1.73cm 0cm]{figures/polarxy_fargo_ex_250}\includegraphics[scale=0.4,clip=true,trim=2.24cm
  0cm 0.0cm 0cm]{figures/polarxy_fargo_ex_300}
  \caption{Evolutoin of the $m=1$ surface density in the FARGO
    simulation with the special $m=1$ initial perturbations
    described in \S\ref{fargo_nonlin}.\label{2d_angmom_ex_evol}} 
\end{figure*}

\begin{figure}
  % scale=0.41
  \includegraphics[width=\linewidth]{figures/nonaxi_evol_ang_fargo_ex}
  \caption{Evolution of angular momentum components in the 
    FARGO simulation with the special $m=1$ initial perturbations
    described in \S\ref{fargo_nonlin}. 
    The perturbation
    relative to $t=0$ in 2D is shown in units of the
    initial total angular momentum $J_\mathrm{ref}$.\label{2d_angmom_ex}} 
\end{figure}
